Text embedding models yield high-resolution insights into conceptual knowledge from short multiple-choice quizzes

### 2023-10-25 paper/main.tex
udpated knowledge map figures

That the knowledge predictions derived from the text embedding space reliably
distinguish between correctly and incorrectly answered held-out questions
(Fig.~\ref{fig:predictions}) suggests that spatial relationships within this
 space can help explain what participants know. But how far does this
explanatory power extend? For example, suppose we know that a participant
correctly answered a question at embedding coordinate $x$. As we move farther away
from $x$ in the embedding space, how does the likelihood that the participant knows about the content at a given location ``fall off'' with distance?
Conversely, suppose the participant instead answered that same question
\textit{in}correctly. Again, as we move farther away from $x$ in the embedding space,
how does the likelihood that the participant does \textit{not} know about a coordinate's
content change with distance? We reasoned that, assuming our embedding space is capturing
something about how individuals actually organize 
their knowledge, a participant's ability to answer questions 
embedded very close to $x$ should tend to be similar to their 
ability to answer the question embedded \textit{at} $x$. Whereas 
at another extreme, once we reach some sufficiently large 
distance from $x$, our ability to infer whether or not a participant 
will correctly answer a question based on their ability to answer 
the question at $x$ should be no better than guessing based on 
their \textit{overall} proportion of correctly answered questions. In 
other words, beyond the maximum distance at which the participant's
 ability to answer the question at $x$ is informative of their ability to 
 answer a second question at location $y$, then guessing the outcome 
 at $y$ based on $x$ should be no more successful than guessing 
 based on a measure that does not consider embedding space distance.

>>>>

That the knowledge predictions derived from the text embedding space reliably
distinguish between held-out correctly versus incorrectly answered questions
(Fig.~\ref{fig:predictions}) suggests that spatial relationships within this
space can help explain what participants know. But how far does this
explanatory power extend? For example, suppose we know that a participant
correctly answered a question at embedding coordinate $x$. As we move farther
away from $x$ in the embedding space, how does the likelihood that the
participant knows about the content at a given location ``fall off'' with
distance? Conversely, suppose the participant instead answered that same
question \textit{in}correctly. Again, as we move farther away from $x$ in the
embedding space, how does the likelihood that the participant does \textit{not}
know about a coordinate's content change with distance? We reasoned that,
assuming our embedding space is capturing something about how individuals
actually organize their knowledge, a participant's ability to answer questions
embedded very close to $x$ should tend to be similar to their ability to answer
the question embedded \textit{at} $x$. Whereas at another extreme, once we
reach some sufficiently large distance from $x$, our ability to infer whether
or not a participant will correctly answer a question based on their ability to
answer the question at $x$ should be no better than guessing based on their
\textit{overall} proportion of correctly answered questions. In other words,
beyond the maximum distance at which the participant's ability to answer the
question at $x$ is informative of their ability to answer a second question at
location $y$, then guessing the outcome at $y$ based on $x$ should be no more
successful than guessing based on a measure that does not consider embedding
space distance.

### 2023-10-25 paper/old2.tex
udpated knowledge map figures

That the knowledge predictions derived from the text embedding space reliably
distinguish between held-out correctly versus incorrectly answered questions
(Fig.~\ref{fig:predictions}) suggests that spatial relationships within this
space can help explain what participants know. But how far does this
explanatory power extend? For example, suppose we know that a participant
correctly answered a question at embedding coordinate $x$. As we move farther
away from $x$ in the embedding space, how does the likelihood that the
participant knows about the content at a given location ``fall off'' with
distance? Conversely, suppose the participant instead answered that same
question \textit{in}correctly. Again, as we move farther away from $x$ in the
embedding space, how does the likelihood that the participant does \textit{not}
know about a coordinate's content change with distance? We reasoned that,
assuming our embedding space is capturing something about how individuals
actually organize their knowledge, a participant's ability to answer questions
embedded very close to $x$ should tend to be similar to their ability to answer
the question embedded \textit{at} $x$. Whereas at another extreme, once we
reach some sufficiently large distance from $x$, our ability to infer whether
or not a participant will correctly answer a question based on their ability to
answer the question at $x$ should be no better than guessing based on their
\textit{overall} proportion of correctly answered questions. In other words,
beyond the maximum distance at which the participant's ability to answer the
question at $x$ is informative of their ability to answer a second question at
location $y$, then guessing the outcome at $y$ based on $x$ should be no more
successful than guessing based on a measure that does not consider embedding
space distance.

>>>>

That the knowledge predictions derived from the text embedding space reliably
distinguish between held-out correctly versus incorrectly answered questions
(Fig.~\ref{fig:predictions}) suggests that spatial relationships in the text
embedding space can help explain what participants know. But how far does this
explanatory power extend? For example, suppose we know that a participant
answers a question (at embedding coordinate $x$) correctly. As we move away
from $x$ in the embedding space, how does their knowledge ``fall off'' with
distance? Or, suppose the participant instead answered that same question
\textit{in}correctly. Again, as we move away from $x$ in the embedding space,
how do the chances that the participant does \textit{not} know about the
content change with distance? We reasoned that, assuming our space is capturing
something about how participants actually organize their knowledge, conceptual
knowledge right around $x$ should be similar to the participant's knowledge of
the content at $x$. And at another extreme, at some distance (after moving
sufficiently far away from $x$), our guesses about what participants know
(based on their response to the question at location $x$) should be no better
than guessing based on their overall proportion of correctly answered
questions—i.e., if $y$ is very far away from $x$, all we can do with the
participant's response to $x$ is guess that ``their performance on quiz
questions about $y$ will be about equal to their average performance on quiz
questions about any material.''
